Structuring College Access: The Market Segment Model and College Board Geomarkets

### 2025-04-21 text/back_end_KS.qmd
adding figs

In contrast, these representational patterns are reversed for Black and Hispanic prospects in @fig-chicago-rq2-lowSAT-race. Among the more than 2,000 Black students included in these orders, the City of Chicago and South and Southwest Suburbs contributed disproportionately higher shares - 48% and 34%, respectively- compared to their overall contributions of 20% and 26% of all included prospects, whereas all other Geomarkets contributed substantially smaller proportions of Black prospects. The City of Chicago Geomarket also contributed larger shares of the nearly 5,000 Hispanic students (35% versus 20% for all prospects), while all other Geomarkets contributed comprable or relatively smaller proportions of Hispanic prospects than their overall representation among all included prospects.

@@

Geomarket contribution patterns to prospect pools by both first-generation college student status and race/ethnicity become more pronounced as SAT score thresholds increase. For instance, @fig-chicago-rq2-midSAT-combo presents prospect pools disaggregated by both first-generation college student status and race/ethnicity within the middle-range SAT score (1160–1300) orders. The City of Chicago Geomarket still contributes a disproportionately larger share of first -generation college students - both those whose parents did not attend college and those whose parents attended some college but did not complete a degree - across most racial and ethnic groups. However, the magnitude of this disproportionately for City of Chicago becomes even greater for Asian students, with a nearly 30-percentage-point difference collectively for all first-generation college students (35% no college and 22% some college) relative to their overall Geomarket representation (28%). Similar to patterns observed at lower SAT score thresholds, the Geomarket contributions to first-generation Hispanic college students remain relatively balanced. All Geomarkets, with the exception of North Shore (22%) and Evanston and Skokie (32%), contributed more than 50% first-generation Hispanic college students across both parental education categories to the overall prospect pool.

@@

The six Geomarkets within the Dallas-Fort Worth metropolitan area are illustrated in @fig-dallas-maps by total population across the three census periods. The Dallas-Fort Worth region represents a unique metroplex case formed by the convergence of two separate metropolitan areas, which creates a unique urban spatial structure that differs from the classic model originally proposed by @RN4982. As depicted in @fig-dallas-maps, two distinct inner-city core Geomarkets - City of Dallas and City of Fort Worth- are seperated by a centrally located suburban Geomarket encompassing Irving, Arlington, & Grand Prairie. These central Geomarkets are surrounded by the more conventional outre ring of suburban Geomarkets (Dallas County, Collin and Rockwall Counties, and West of Dallas/Ft. Worth Metroplex).

@@

Figure X presents the racial and ethnic composition of each Geomarket within the Dallas-Fort Worth metropolitan area across the three census periods. In 1980, all Geomarkets were predominantly White, with White populations ranging from 62% in City of Dallas to nearly 92% in West of Dallas/Ft. Worth Metroplex. By 2000, however, Geomarkets located in the central part of the metropolitan area exhibited a a more balanced racial/ethnic distribution, including Dallas County (58% White, 16% Black, 19% Hispanic; 6% Asian Pacific Islander), Irving, Arlington, and Grand Prairie (54% White, 12% Black, 26% Hispanic; 16% Asian Pacific Islander), and City of Fort Worth (57% White, 16% Black, 22% Hispanic; 3% Asian Pacific Islander). In contrast, Collin & Rockwall Counties and West of Dallas/Fort Worth Metroplex Geomarkets remained predominantly White in 2000 - 78% and 80%, respectively. Notably, by 2000, the City of Dallas was no longer a predominantly White Geomarket, with a racial/ethnic composition of approximately 24% Black, 30% Hispanic, and 3% Asian Pacific Islander residents.

@@

@@

```{r fig-chicago-rq2-lowSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, Low-Range SAT orders", out.width = "100%"}
knitr::include_graphics('./../assets/images/rq2_chicago_lowSAT_race.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```

@@

```{r fig-chicago-rq2-highSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, High-Range SAT orders", out.width = "100%"}
knitr::include_graphics('./../assets/images/rq2_chicago_highSAT_race.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```

```{r fig-chicago-rq2-lowSAT-ses, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, Low-Range SAT orders", out.width = "100%"}
knitr::include_graphics('./../assets/images/rq2_chicago_lowSAT_ses.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```

@@

```{r fig-chicago-rq2-highSAT-ses, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, High-Range SAT orders", out.width = "100%"}
knitr::include_graphics('./../assets/images/rq2_chicago_highSAT_ses.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```

```{r fig-chicago-rq2-lowSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, Low-Range SAT orders", out.width = "100%"}
knitr::include_graphics('./../assets/images/rq2_chicago_lowSAT_combo.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```

@@

```{r fig-chicago-rq2-highSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, High-Range SAT orders", out.width = "100%"}
knitr::include_graphics('./../assets/images/rq2_chicago_highSAT_combo.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png
```

@@

>>>>

In contrast, these representational patterns are reversed for Black and Hispanic prospects in @fig-chicago-rq2-lowSAT-race. Among the more than 2,000 Black students included in these orders, the City of Chicago and South and Southwest Suburbs contributed disproportionately higher shares, 48% and 34%, respectively, compared to their overall contributions of 20% and 26% of all included prospects, whereas all other Geomarkets contributed substantially smaller proportions of Black prospects. The City of Chicago Geomarket also contributed larger shares of the nearly 5,000 Hispanic students (35% versus 20% for all prospects), while all other Geomarkets contributed comprable or relatively smaller proportions of Hispanic prospects than their overall representation among all included prospects.

@@

Geomarket contribution patterns to prospect pools by both first-generation college student status and race/ethnicity become more pronounced as SAT score thresholds increase. For instance, @fig-chicago-rq2-midSAT-combo presents prospect pools disaggregated by both first-generation college student status and race/ethnicity within the middle-range SAT score (1160–1300) orders. The City of Chicago Geomarket still contributes a disproportionately larger share of first-generation college students for both those whose parents did not attend college and those whose parents attended some college but did not complete a degree across most racial and ethnic groups. However, the magnitude of this disproportionately for City of Chicago becomes even greater for Asian students, with a nearly 30-percentage-point difference collectively for all first-generation college students (35% no college and 22% some college) relative to their overall Geomarket representation (28%). Similar to patterns observed at lower SAT score thresholds, the Geomarket contributions to first-generation Hispanic college students remain relatively balanced. All Geomarkets, with the exception of North Shore (22%) and Evanston and Skokie (32%), contributed more than 50% first-generation Hispanic college students across both parental education categories to the overall prospect pool.

@@

The six Geomarkets within the Dallas-Fort Worth metropolitan area are illustrated in @fig-dallas-maps by total population across the three census periods. The Dallas-Fort Worth region represents a unique metroplex case formed by the convergence of two separate metropolitan areas, which creates a unique urban spatial structure that differs from the classic model originally proposed by @RN4982. As depicted in @fig-dallas-maps, two distinct inner-city core Geomarkets (City of Dallas and City of Fort Worth) are seperated by a centrally located suburban Geomarket encompassing Irving, Arlington, & Grand Prairie. These central Geomarkets are surrounded by the more conventional outre ring of suburban Geomarkets (Dallas County, Collin and Rockwall Counties, and West of Dallas/Ft. Worth Metroplex).

@@

Figure X presents the racial and ethnic composition of each Geomarket within the Dallas-Fort Worth metropolitan area across the three census periods. In 1980, all Geomarkets were predominantly White, with White populations ranging from 62% in City of Dallas to nearly 92% in West of Dallas/Ft. Worth Metroplex. By 2000, however, Geomarkets located in the central part of the metropolitan area exhibited a a more balanced racial/ethnic distribution, including Dallas County (58% White, 16% Black, 19% Hispanic; 6% Asian Pacific Islander), Irving, Arlington, and Grand Prairie (54% White, 12% Black, 26% Hispanic; 16% Asian Pacific Islander), and City of Fort Worth (57% White, 16% Black, 22% Hispanic; 3% Asian Pacific Islander). In contrast, Collin & Rockwall Counties and West of Dallas/Fort Worth Metroplex Geomarkets remained predominantly White in 2000 at 78% and 80%, respectively. Notably, by 2000, the City of Dallas was no longer a predominantly White Geomarket, with a racial/ethnic composition of approximately 24% Black, 30% Hispanic, and 3% Asian Pacific Islander residents.

@@

The 11 Geomarkets within the Los Angeles metropolitan area are illustrated in Figure X by total population across the three census periods. The Los Angeles region is one of the only multi-nuclei cases metroplex case formed by the convergence of two separate metropolitan areas, which creates a unique urban spatial structure that differs from the classic model originally proposed by @RN4982. As depicted in @fig-dallas-maps, two distinct inner-city core Geomarkets - City of Dallas and City of Fort Worth- are seperated by a centrally located suburban Geomarket encompassing Irving, Arlington, & Grand Prairie. These central Geomarkets are surrounded by the more conventional outre ring of suburban Geomarkets (Dallas County, Collin and Rockwall Counties, and West of Dallas/Ft. Worth Metroplex).

@@

```{r fig-chicago-rq2-race, echo = FALSE, fig.align = 'center', fig.cap = "Racial/Ethnic Composition of Purchased Student Profiles by Geomarket, Chicago Area", out.width = "100%"}
knitr::include_graphics('./../assets/images/rq2_chicago_race.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```

@@

@@

@@

@@

```{r fig-dallas-rq1-race, echo = FALSE, fig.align = 'center', fig.cap = "Racial/Ethnic Composition of Dallas Area Geomarkets, 1980-2020", out.width = "100%"}
knitr::include_graphics('./../assets/images/rq1-dallas-race.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```
